### 2023-08-15 proofs/tex/wf-lemmas.tex
WIP: leftovers

\begin{lemma}[Equivalent Contexts]
  \label{lemma:wf-ctxt-equiv}

  In the well-formedness judgment, only used variables matter:
  \begin{itemize}
  \item[$+$] if $[[{Γ1} ∩ fv iP]] = [[{Γ2} ∩ fv iP]]$ then
    $[[Γ1 ⊢ iP]] \iff [[Γ2 ⊢ iP]]$,
  \item[$-$] if $[[{Γ1} ∩ fv iN]] = [[{Γ2} ∩ fv iN]]$ then
    $[[Γ1 ⊢ iN]] \iff [[Γ2 ⊢ iN]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  By simple mutual induction on $[[iP]]$ and $[[iQ]]$. 
\end{proof}

\begin{lemma}[Well-formedness Context Weakening]
  \label{lemma:wf-weakening}
  Suppose that $[[{Γ1} ⊆ {Γ2}]]$, then
  \begin{itemize}
    \item[$+$] if $[[Γ1 ⊢ iP]]$ then $[[Γ2 ⊢ iP]]$,
    \item[$-$] if $[[Γ1 ⊢ iN]]$ then $[[Γ2 ⊢ iN]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  By simple mutual induction on $[[iP]]$ and $[[iQ]]$. 
\end{proof}

>>>>

\begin{lemma}[Soundness of well-formedness]
    \label{lemma:wf-soundness}
    \hfill

  \begin{itemize}
    \item[$+$] if $[[Γ ⊢ iP]]$ then $[[fv(iP) ⊆ {Γ}]]$;
    \item[$-$] if $[[Γ ⊢ iN]]$ then $[[fv(iN) ⊆ {Γ}]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  The proof is done by a simple structural induction on 
  $[[Γ⊢ iP]]$ and mutually, $[[Γ ⊢ iN]]$.
  \begin{caseof}
    \item $[[Γ ⊢ α±]]$ means by inversion that $[[α±]] \in [[Γ]]$,
      that is, $[[α±]] = [[fv(α±) ⊆ {Γ}]]$.
    \item $[[Γ ⊢ iQ → iM]]$ means by inversion that
      $[[Γ ⊢ iQ]]$ and $[[Γ ⊢ iM]]$.
      Then by the induction hypothesis, $[[fv(iQ) ⊆ {Γ}]]$ and $[[fv(iM) ⊆ {Γ}]]$,
      and hence, $[[fv(iQ → iM) = fv(iQ) ∪ fv(iM)]] \subseteq [[{Γ}]]$.
    \item the cases when $[[iP = ↓iN']]$ or $[[iN = ↑iP']]$ are proven analogously.
    \item $[[Γ ⊢ ∀pas.iM]]$ means by inversion that $[[Γ, pas ⊢ iM]]$.
      Then by the induction hypothesis, $[[fv(iM) ⊆ {Γ, pas}]]$, and hence,
      $[[fv(∀pas.iM) = fv(iM) \ {pas}]] \subseteq [[{Γ, pas} \ {pas} = {Γ}]]$.
    \item The case $[[iP = ∃nas.iQ]]$ is proven analogously.
  \end{caseof}
\end{proof}

\begin{lemma}[Completeness of well-formedness]
  \label{lemma:wf-ctxt-equiv}

  In the well-formedness judgment, only used variables matter:
  \begin{itemize}
  \item[$+$] if $[[{Γ1} ∩ fv iP]] = [[{Γ2} ∩ fv iP]]$ then
    $[[Γ1 ⊢ iP]] \iff [[Γ2 ⊢ iP]]$,
  \item[$-$] if $[[{Γ1} ∩ fv iN]] = [[{Γ2} ∩ fv iN]]$ then
    $[[Γ1 ⊢ iN]] \iff [[Γ2 ⊢ iN]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  By simple mutual induction on $[[iP]]$ and $[[iN]]$.
\end{proof}

\begin{corollary}[Context Strengthening]
  \label{corollary:wf-ctxt-strengthening}
  \hfill

  \begin{itemize}
    \item [$+$] If $[[Γ ⊢ iP]]$ then $[[Ord fv(iP) ⊢ iP]]$;
    \item [$-$] If $[[Γ ⊢ iN]]$ then $[[Ord fv(iN) ⊢ iN]]$.
  \end{itemize}
\end{corollary}
\begin{proof}
  It follows from \cref{lemma:wf-ctxt-equiv} and \cref{lemma:wf-soundness}.
  \begin{itemize}
    \item [$+$] 
      By \cref{lemma:wf-soundness}, $[[fv(iP) ⊆ {Γ}]]$, and hence,
      $[[{Γ} ∩ fv iP = fv iP ]]$, which makes \cref{lemma:wf-ctxt-equiv}
      applicable fore contexts $[[Γ]]$ and $[[Ord fv(iP)]]$.
    \item[$-$] The negative case is proven analogously.
  \end{itemize}

\begin{corollary}[Well-formedness Context Weakening]
  \label{lemma:wf-weakening}
  Suppose that $[[{Γ1} ⊆ {Γ2}]]$, then
  \begin{itemize}
    \item[$+$] if $[[Γ1 ⊢ iP]]$ then $[[Γ2 ⊢ iP]]$,
    \item[$-$] if $[[Γ1 ⊢ iN]]$ then $[[Γ2 ⊢ iN]]$.
  \end{itemize}
\end{corollary}
\begin{proof}
  By \cref{lemma:wf-soundness},
  $[[Γ1 ⊢ iP]]$ implies $[[fv(iP) ⊆ {Γ1}]]$,
  which means that $[[fv(iP) ⊆ {Γ2}]]$,
  and thus, $[[fv(iP)]] = [[fv(iP) ∩ {Γ1}]] = [[fv(iP) ∩ {Γ2}]]$.
  Then by \cref{lemma:wf-ctxt-equiv}, $[[Γ2 ⊢ iP]]$. 
  The negative case is symmetric.
\end{proof}
